---
title: "Banking Capstone - Data Wrangling and Stastical Analysis"
author: "Demetri Lee"
date: "December 1, 2018"
geometry: margin=1in
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(GGally)
library(summarytools)
knitr::opts_chunk$set(echo = TRUE)
```

## Working to predict the success of bank marketing

I'm working on predicting the success of telemarketing calls for a bank that's 
selling long-term deposits. A Portuguese retail bank was addressed, with data 
collected from 2008 to 2013, thus including the effects of the recent 
financial crisis of 2008. With 21 variables available for analysis, I hope to 
derive knowledge for a model that would confirm how to better focus resources 
towards clients with a high chance of agreeing to registering for extra 
banking services. 

### This is the data dictionary provided by the respective retail bank:

**Bank client data**

1. age: numeric
2. job: type of job (categorical: admin, blue-collar, entrepreneur,
housemaid, management, retired, self-employed, services, student,
technician, unemployed, unknown)
3. marital: marital status (categorical: divorced, married, single,
unknown; note: 'divorced' means divorced or widowed)
4. education: (categorical: basic.4y, basic.6y, basic.9y, high.school,
illiterate, professional.course, university.degree, unknown)
5. default: has credit in default? (categorical: no, yes, unknown)
6. housing: has housing loan? (categorical: no, yes, unknown)
7. loan: has personal loan? (categorical: no, yes, unknown)

**Related with the last contact of the current campaign**

8. contact: contact communication type (categorical: cellular, telephone) 
9. month: last contact month of year (categorical: jan', 'feb', 'mar', ..., 
'nov', 'dec')
10. day_of_week: last contact day of the week (categorical: mon, tue, wed,
thu, fri)
11. duration: last contact duration, in seconds (numeric). 
Important note: this attribute highly affects the output target (e.g., if 
duration=0 then y='no'). Yet, the duration is not known before a call is 
performed. Also, after the end of the call y is obviously known. Thus, this 
input should only be included for benchmark purposes and should be discarded 
if the intention is to have a realistic predictive model.

**Other attributes**

12. campaign: number of contacts performed during this campaign and for this 
client (numeric, includes last contact)
13. pdays: number of days that passed by after the client was last contacted 
from a previous campaign (numeric; 999 means client was not previously contacted)
14. previous: number of contacts performed before this campaign and for this 
client (numeric)
15. poutcome: outcome of the previous marketing campaign (categorical: 
failure, nonexistent, success)

**Social and economic context attributes**

16. emp.var.rate: employment variation rate - quarterly indicator (numeric)
17. cons.price.idx: consumer price index - monthly indicator (numeric) 
18. cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
19. euribor3m: euribor 3 month rate - daily indicator (numeric)
20. nr.employed: number of employees - quarterly indicator (numeric)

**Output variable (desired target)**

21. y: has the client subscribed a term deposit? (binary: yes, no)

### Data Preparation

Let's start by reading in the dataset with values delimited by semi-colons along 
with designated column types as indicated by the data dictionary.
```{r Data_Prep}
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "ifffffffffiiiifdddddf")
summary(bankfull)
```

Let's rename variables to be more descriptive and avoid use of periods.
```{r}
bankfull <- rename(bankfull, edu_lvl = "education")
bankfull <- rename(bankfull, cred_default = "default")
bankfull <- rename(bankfull, mortgage = "housing")
bankfull <- rename(bankfull, duration_sec = "duration")
bankfull <- rename(bankfull, contact_typ = "contact")
bankfull <- rename(bankfull, contact_cnt = "campaign")
bankfull <- rename(bankfull, days_passed = "pdays")
bankfull <- rename(bankfull, prev_contact_cnt = "previous")
bankfull <- rename(bankfull, prev_outcome = "poutcome")
bankfull <- rename(bankfull, emp_var_rate = "emp.var.rate")
bankfull <- rename(bankfull, cons_price_idx = "cons.price.idx")
bankfull <- rename(bankfull, cons_conf_idx = "cons.conf.idx")
bankfull <- rename(bankfull, num_employed = "nr.employed")
```

The goal is to determine which factors lead to a subscription so I'll rename 
the dependent variable to `subscribed`, which is more descriptive than just `y`.
```{r}
bankfull <- rename(bankfull, subscribed = y)
```

As per data dictionary, if `duration` equals 0 then customers is automatically 
not subscribed. Let's create a new variable, `contacted` for results.
```{r}
bankfull <- mutate(bankfull, contacted = (duration_sec != "0"))
bankfull$contacted <- as.factor(bankfull$contacted)
```

Replace all periods ('.') in values with underscores ('_') for legibility.
```{r}
summary(bankfull$job) # job types before replacement 
jtemp <- bankfull$job
bankfull$job <- gsub("\\.", "", jtemp)
bankfull$job <- as.factor(bankfull$job)
summary(bankfull$job) # job types after replacement

summary(bankfull$edu_lvl) # education levels before replacement
etemp <- bankfull$edu_lvl
bankfull$edu_lvl <- gsub("\\.", "_", etemp)
bankfull$edu_lvl <- as.factor(bankfull$edu_lvl)
summary(bankfull$edu_lvl) # education levels after replacement
```

Take a quick check of the dataset with glimpse.
```{r}
glimpse(bankfull)
```

### Explorating Data with Plots

**Looking at Campaign results**

A cursory look at the `subscribe` variable will show most customers do not sign 
up for the extra service.
```{r}
ggplot(bankfull, aes(x = subscribed)) + geom_bar() + 
  labs(title = "Subscription Counts")
```

The proportion of `no` to `yes` in a table:
```{r}
prop.table(table(bankfull$subscribed))
#bankfull %>% knitr::kable(digits = 2)
```

When looking for clients that have subscribed to a previous campaign there are 
too many `nonexistent` values. 
```{r}
ggplot(bankfull, aes(prev_outcome)) + geom_bar() + 
  labs(title="Subscription Count from a Previous Campaign")
```

When comparing clients that subscribed to the previous campaign and the 
current campaign, the `nonexistent` values become distracting.
```{r}
ggplot(bankfull, aes(x=prev_outcome, fill=subscribed)) + geom_bar()
```

If the `nonexistent` values are removed, more than half the clients that 
subscribed previously will subscribe again.
```{r}
bankfull_prev <- filter(bankfull, prev_outcome != "nonexistent")
ggplot(bankfull_prev, aes(x=prev_outcome, fill=subscribed)) + 
  geom_bar(position="fill") + 
  labs(title = "Proportion of subscriptions from previous campaign")

prop.table(table(bankfull_prev$prev_outcome))
```


**Looking at Client Data**

Let's take a look at whether there is any correlation with client data.
```{r}
ggplot(bankfull, aes(x = age, fill = subscribed)) + 
  geom_bar() + 
  labs(title = "Subscriptions per Age Group")

ggplot(bankfull, aes(x = edu_lvl, fill = subscribed)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Subscriptions per Education Level")

ggplot(bankfull, aes(x = job, fill = subscribed)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Subscriptions per Job Type") # avoid using Ques in title



```

Density of Customer Age
```{r}
plot(density(bankfull$age), main="Density of Customer Age")
ggplot(bankfull, aes(x=age)) + geom_bar()
ggplot(bankfull, aes(x=age)) + geom_histogram(binwidth = 1)
```

Checking elsewhere, is there any positive correlation between those who 
`subscribe` and have a default on credit, mortgage, or a personal loan?
```{r}

ggpairs(bankfull, columns=5:7)


ggplot(bankfull, aes(x = cred_default, fill = subscribed)) + 
  geom_bar(position="fill") + 
  labs(title = "Subscriptions relative to Clients who Default on Credit")

# -- Swap the x and fill fields -- #
ggplot(bankfull, aes(x = mortgage, fill = subscribed)) + 
  geom_bar(position="fill") + 
  labs(title = "Subscriptions relative to Clients with a Mortgage")

ggplot(bankfull, aes(x = loan, fill = subscribed)) + 
  geom_bar(position="fill") + 
  labs(title = "Susbscriptions relative to Clients with an Active Loan")

#ggplot(bankfull, aes(x = mortgage, fill = subscribed)) + geom_bar(position = "fill") + facet_wrap(~loan)

#3 x 3 pair plot > compared with "subscribed"

```

_It seems those with defaults on credit or personal loans are not lightly to 
subscribe but those with a housing loan may subscribe._

**Last Contact Data**

The variable _duration_ refers to the length of time the customer was last 
contacted, in seconds. Important note:  this attribute highly affects the 
output target (e.g., if duration=0 then subscribed="no"). Yet, the duration 
is not known before a call is performed. Also, after the end of the call 
_subscribed_ is obviously known. Thus, this input should only be included 
for benchmark purposes and should be discarded if the intention is to have a 
realistic predictive model.

```{r}
bankfull <- bankfull %>% mutate(duration_min = duration_sec/60)
summary(bankfull$duration_min)
ggplot(bankfull, aes(x=duration_min, fill=subscribed)) + geom_bar()

#time_to_subscribed <- bankfull %>% filter(subscribed)
ggplot(bankfull, aes(x=subscribed, y=duration_min)) + 
  geom_boxplot(outlier.colour="blue", outlier.shape=6, outlier.size=2)
ggplot(bankfull, aes(x=subscribed, y=duration_min)) + 
  geom_boxplot() + coord_flip()
```

It appears customers are more inclined to subscribe when contacted via 
cellular compared to telephone.
```{r}
ggplot(bankfull, aes(x=contact_typ, fill = subscribed)) + geom_bar()
ggplot(bankfull, aes(x=contact_typ, fill = subscribed)) + 
  geom_bar(position="fill") + 
  labs(title="Contact by Cellular or Telephone")

#ggplot(bankfull, aes(x=contact_typ, y=age)) + 
#  geom_col() + 
#  labs(title="Contact by Cellular or Telephone")

table(bankfull$contact_typ)
```

A regular bar graph by count makes it difficult to determine which month a 
client is most likely to subscribe. At best it looks like *May* has the most 
subscribers, but that doesn't mean it's the most likely.
```{r}
ggplot(bankfull, aes(x=month, fill=subscribed)) + geom_bar()
```

But if we stack each month as relative proportions we can see customers are more receptive to subscribing in *March* followed by *December* before *September*.
```{r}
ggplot(bankfull, aes(x=month, fill=subscribed)) + geom_bar(position="fill")
```

Updated to include weekdays and we see the second column, *Wednesday*, becomes a good day to subscribe.
```{r}
ggplot(bankfull, aes(x=day_of_week, fill=subscribed)) + 
  geom_bar(position="fill") +
  facet_wrap(vars(month))
```

The most subscriptions actually happen in *May* and other summer months, 
*June, July, and August*.
```{r}

monthsCampaign <- bankfull %>% filter(subscribed == "yes") %>% select(month)
freq(monthsCampaign)

```


**Socioeconomic Data**

There looks to be a high correlation in the following 2 pairs:

* euribor3m vs employment variation rate
* euribor3m vs number employed at bank 
```{r}
# From emp_var_rate to num_employed
ggpairs(bankfull, columns=16:19)

```

```{r}
ggplot(bankfull, aes(emp_var_rate, euribor3m)) + geom_col()
ggplot(bankfull, aes(euribor3m, emp_var_rate)) + geom_col()
ggplot(bankfull, aes(num_employed, euribor3m)) + geom_col()
ggplot(bankfull, aes(euribor3m, num_employed)) + geom_col()
```


_Posit questions as they come up from observations_

1. Can you count something interesting?
2. Can you find some trends (high, low, increase, decrease, anomalies)?
3. Can you make a bar plot or a histogram?
```{r}
ggplot(bankfull, aes(x=duration_sec)) + geom_histogram(bins=10) + facet_wrap(~subscribed)
ggplot(bankfull, aes(x=duration_sec)) + geom_histogram(bins=10) + facet_wrap(~prev_outcome)
```


4. Can you compare two related quantities?
```{r}
ggplot(bankfull, aes(x=subscribed, y=emp_var_rate)) + geom_boxplot()
ggplot(bankfull, aes(x=subscribed, y=euribor3m)) + geom_boxplot()
```

5. Can you make a scatterplot?
```{r}
ggplot(bankfull, aes(x=emp_var_rate, y=euribor3m)) + 
  geom_point(aes(colour = factor(subscribed)), alpha = 0.50, position="jitter") + 
  labs(title = "Euribor (3 months) vs Employment Variation Rate (quarterly)")

```

6. Can you make a time-series plot?

Write the revised dataset into a cleaned-up file.
```{r}
write_csv(bankfull, path="bankfull_clean.csv")
```

