---
title: "Banking Capstone - Data Wrangling and Stastical Analysis"
author: "Demetri Lee"
date: "December 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Trying to predict the success of bank telemarketing

I'm working on predicting the success of telemarketing calls for a bank that's 
selling long-term deposits. A Portuguese retail bank was addressed, with data 
collected from 2008 to 2013, thus including the effects of the recent 
financial crisis of 2008. With 21 features available for analysis, I hope to 
derive knowledge for a model that would confirm how to better focus resources 
towards clients with a high chance of agreeing to registering for extra 
banking services. 

### This is the data dictionary provided by the respective retail bank:

**Bank client data**

1. age, numeric
2. job: type of job (categorical: admin, blue-collar, entrepreneur,
housemaid, management, retired, self-employed, services, student,
technician, unemployed, unknown)
3. marital: marital status (categorical: divorced, married, single,
unknown; note: 'divorced' means divorced or widowed)
4. education: categorical (basic.4y, basic.6y, basic.9y, high.school,
illiterate, professional.course, university.degree, unknown)
5. default: has credit in default? (categorical: no, yes, unknown)
6. housing: has housing loan? (categorical: no, yes, unknown)
7. loan: has personal loan? (categorical: no, yes, unknown)

**Related with the last contact of the current campaign**

8. contact: contact communication type (categorical: cellular, telephone) 
9. month: last contact month of year (categorical: jan', 'feb', 'mar', ..., 
'nov', 'dec')
10. day_of_week: last contact day of the week (categorical: mon, tue, wed,
thu, fri)
11. duration: last contact duration, in seconds (numeric). 
Important note: this attribute highly affects the output target (e.g., if 
duration=0 then y='no). Yet, the duration is not known before a call is 
performed. Also, after the end of the call y is obviously known. Thus, this 
input should only be included for benchmark purposes and should be discarded 
if the intention is to have a realistic predictive model.

**Other attributes**

12. campaign: number of contacts performed during this campaign and for this 
client (numeric, includes last contact)
13. pdays: number of days that passed by after the client was last contacted 
from a previous campaign (numeric; 999 means client was not previously contacted)
14. previous: number of contacts performed before this campaign and for this 
client (numeric)
15. poutcome: outcome of the previous marketing campaign (categorical: 
failure, nonexistent, success)

**Social and economic context attributes**

16. emp.var.rate: employment variation rate - quarterly indicator (numeric)
17. cons.price.idx: consumer price index - monthly indicator (numeric) 
18. cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
19. euribor3m: euribor 3 month rate - daily indicator (numeric)
20. nr.employed: number of employees - quarterly indicator (numeric)

**Output variable (desired target)**

21. y: has the client subscribed a term deposit? (binary: yes, no)

**Read in the dataset with values delimited by semi-colons along with 
designated column types as indicated by the data dictionary.**
```{r}
#bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
 bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "ifffffffffiiiicdddddc")
```

Let's rename variables to be more descriptive and avoid use of periods (".")
```{r}
bankfull <- rename(bankfull, edu_lvl = "education")
bankfull <- rename(bankfull, cred_default = "default")
bankfull <- rename(bankfull, mortgage = "housing")
bankfull <- rename(bankfull, duration_sec = "duration")
bankfull <- rename(bankfull, contact_typ = "contact")
bankfull <- rename(bankfull, contact_cnt = "campaign")
bankfull <- rename(bankfull, days_passed = "pdays")
bankfull <- rename(bankfull, prev_contact_cnt = "previous")
bankfull <- rename(bankfull, prev_outcome = "poutcome")
bankfull <- rename(bankfull, emp_var_rate = "emp.var.rate")
bankfull <- rename(bankfull, cons_price_idx = "cons.price.idx")
bankfull <- rename(bankfull, cons_conf_idx = "cons.conf.idx")
bankfull <- rename(bankfull, num_employed = "nr.employed")
```

As per data dictionary, if `duration`equals 0 then customers is automatically 
not subscribed. Let's create a new feature/variable, `contacted` for results.
```{r}
bankfull <- mutate(bankfull, contacted = (duration_sec != "0"))
```


The goal is to determine which factors lead to a subscription so I'll rename 
the dependent variable to `subscribed`, which is more descriptive than just `y`.
```{r}
bankfull <- rename(bankfull, subscribed = y)
```

Replace all periods ('.') with underscores ('_') for legibility.
Also change the _missing values_ to NA.
```{r}
jtemp <- bankfull$job
bankfull$job <- gsub("\\.", "", jtemp)
bankfull$job[bankfull$job == "unknown"] <- NA
bankfull$job <- as.factor(bankfull$job)

etemp <- bankfull$edu_lvl
bankfull$edu_lvl <- gsub("\\.", "_", etemp)
bankfull$edu_lvl[bankfull$edu_lvl == "unknown"] <- NA
bankfull$edu_lvl <- as.factor(bankfull$edu_lvl)
```

There are only 2 possible values for the dependent variable, "subscribed". As
such I'll convert the variable to type factor.
```{r}
bankfull$subscribed <- as.factor(bankfull$subscribed)
```



Take a quick check of the dataset with summary and glimpse.
```{r}
glimpse(bankfull)
summary(bankfull)
```


Let's take care of _missing values_ as necessary
```{r}
bankfull$marital[bankfull$marital == "unknown"] <- NA
bankfull$marital <- as.factor(bankfull$marital)

bankfull$cred_default[bankfull$cred_default == "unknown"] <- NA
bankfull$cred_default <- as.factor(bankfull$cred_default)

bankfull$housing[bankfull$housing == "unknown"] <- NA
bankfull$housing <- as.factor(bankfull$housing)

bankfull$prev_outcome[bankfull$prev_outcome == "nonexistent"] <- NA
bankfull$prev_outcome <- as.factor(bankfull$prev_outcome)
```


### Including Plots

A cursory look at the `subscribe` variable will show most customers do not sign 
up for the extra service.
```{r}
prop.table(table(bankfull$subscribed))
ggplot(bankfull, aes(x = subscribed)) + geom_bar() + labs(title = "How many clients subscribed?")
```

How many clients from the previous campaign also signed up for a term deposit 
this time around? This should be similar to checking for a repeat customer.
```{r}
xtabs(~prev_outcome+subscribed, bankfull)
ggplot(bankfull, aes(prev_outcome)) + geom_bar() + labs(title = "How many clients subscribed to a previous campaign?")
plot(xtabs(~prev_outcome+subscribed, bankfull), 
     main="Previous Outcome vs Current Subscrption")
```

Unfortunately there is a large number of "nonexistent" values for `prev_outome`.

**Although most of the inside sales from the previous campaign resulted in 
failed attempts, the large number of NA results makes this feature useless.**
```{r}
ggplot(bankfull, aes(x = subscribed, fill = prev_outcome)) + 
  geom_bar() + 
  labs(title = "How many customers from previous campaign also bought into current subscription?")
```

**Checking elsewhere, is there any positive correlation between those who 
`subscribe` and have a default on credit, mortgage, or a personal loan?** 
```{r}
ggplot(bankfull, aes(x = subscribed, fill = cred_default)) + 
  geom_bar() + 
  labs(title = "Do customers with credit in default tend to describe?")

ggplot(bankfull, aes(x = subscribed, fill = mortgage)) + 
  geom_bar() + 
  labs(title = "Do customers with a mortgage tend to subscribe?")

ggplot(bankfull, aes(x = subscribed, fill = loan)) + 
  geom_bar() + 
  labs(title = "Do customers with a personal loan tend to subscribe?")

ggplot(bankfull, aes(x = mortgage, fill = subscribed)) + 
  geom_bar() + 
  facet_wrap(~loan)

```

_It seems those with defaults on credit or personal loans are not lightly to 
subscribe but those with a housing loan may subscribe._

**Let's look at the age of our customers**

```{r}
ggplot(bankfull, aes(x = age, fill = subscribed)) + 
  geom_histogram(binwidth = 5) + 
  labs(title = "Age distrubtion of current bank clients and those who subscribed") 
```

```{r}
ggplot(bankfull, aes(x = subscribed, y = age)) + 
  geom_boxplot() + 
  labs(title = "Subscription rates by age")
```



The variable _duration_ refers to the length of time the customer was last 
contacted, in seconds. Important note:  this attribute highly affects the 
output target (e.g., if duration=0 then subscribed="no"). Yet, the duration 
is not known before a call is performed. Also, after the end of the call 
_subscribed_ is obviously known. Thus, this input should only be included 
for benchmark purposes and should be discarded if the intention is to have a 
realistic predictive model.

```{r}
bankfull <- bankfull %>% mutate(duration_min = duration_sec/60)
summary(bankfull$duration_min)
boxplot(bankfull$duration_min)
ggplot(bankfull, aes(x=duration_min)) + geom_bar()
```


```{r}
plot(density(bankfull$age), main="Density of Customer Age")
ggplot(bankfull, aes(x=age)) + geom_bar()
ggplot(bankfull, aes(x=age)) + geom_histogram(binwidth = 1)
```

```{r fig.cap = "Contact Type vs Subscribed"}
#count(bankfull, vars = day_of_week)
#ggplot(bankfull, aes(x=subscribed, fill = contact_typ)) + geom_bar()
ggplot(bankfull, aes(x=contact_typ, fill = subscribed)) + geom_bar()
#try geom_col()
count(bankfull)
```

Outcome of previous campaign

```{r}
ggplot(bankfull, aes(x = prev_outcome)) + geom_bar()
summary(bankfull$prev_outcome)
```


How many of the same customers subscribed to a long-term deposit service versus those who have not?

```{r echo=TRUE, fig.cap = "Only 11.3% of customers subscribed while an overwhelming 88.7% customers declined"}
reject_cnt <- tally(bankfull, subscribed == "no")
reject_pct <- reject_cnt/nrow(bankfull)*100
accept_cnt <- tally(bankfull, subscribed == "yes")
accept_pct <- accept_cnt/nrow(bankfull)*100
summary(bankfull$subscribed)
ggplot(bankfull, aes(x = subscribed)) + geom_bar()

print(accept_pct)
#print(reject_pct, "% would reject.")

```
test = `r reject_cnt + 20`

How did a previous campaign perform?

```{r fig.cap = "failure = 10.3% vs success = 3.3% while remaining 35,563 were non-existent"}
ggplot(bankfull, aes(x = prev_outcome)) + geom_bar()
```

_Posit questions as they come up from observations_

1. Can you count something interesting?
2. Can you find some trends (high, low, increase, decrease, anomalies)?
3. Can you make a bar plot or a histogram?
4. Can you compare two related quantities?
5. Can you make a scatterplot?
6. Can you make a time-series plot?

Write the revised dataset into a cleaned-up file.
```{r}
write_csv(bankfull, path="bankfull_clean.csv")
```

