---
title: "Bank Marketing Success"
author: "Demetri Lee"
date: "09-APR-2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(tidyverse)
library(summarytools)

#bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "ifffffffffiiiifdddddf")

colnames(bankfull) <- c("age", "job", "marital", "edu_lvl", "cred_default", "mortgage", "loan",
                        "contact_typ", "month", "day_of_week", "duration_sec", 
                        "contact_cnt", "days_passed", "prev_contact_cnt", "prev_outcome",
                        "emp_var_rate", "cons_price_idx", "cons_conf_idx", "euribor3m", "num_employed",
                        "subscribed")
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

A Portuguese bank uses telemarketing campaigns to engage their clients.

Each client engagement has been recorded for evaluation.

Such data can be used to extract insight for making better campaign decisions.


## Objective

Marketing campaigns generally do not discriminite which clients are targeted.

As such, unneccessary resources may be spent on clients that lead to rejection. 

The goal is to predict whether a client will subscribe to a term deposit.


## Data

Interactions for over 41,000 clients were recorded from May-2008 to Nov-2010.

Each record contains 21 different attributes for evaluation.

The attribute of interest, `subscribed` will indicate a successful subscription from a client.


## Exploratory Data Analysis

**At first glance there is an overwhelming amount of clients who choose not to 
subscribe to a term deposit.**

```{r subscribed}
ggplot(bankfull, aes(x = subscribed, fill = subscribed)) + geom_bar() + 
  xlab("subscribed") + 
  ylab("count")
```


## Exploratory Data Analysis

**But looking at subscriptions from a previous campaign tells us more than 
half the subscribers will re-subscribe.**

```{r prev_outcome}
bankfull_prev <- filter(bankfull, prev_outcome != "nonexistent")
ggplot(bankfull_prev, aes(x=prev_outcome, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("previous campaign") + 
  ylab("ratio")
```


## Contact Type

**It was also observed that contacting clients via cellular phone yields 2 times more success than reaching out to a client's landline telephone.**

```{r contact_typ}
ggplot(bankfull, aes(x=contact_typ, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("method of contact") + 
  ylab("count")
```


## Employment Variation Rate

**Logistic Regression indicates there is fair chance of success when the 
Employment Variation Rate is -1.8.**

```{r emp_var_rate}
ggplot(bankfull, aes(x=subscribed, y=emp_var_rate, fill=subscribed)) +
  geom_boxplot() + geom_jitter(width=0.2) + 
  xlab("subscribed")
  ylab("employment variation rate")
```


## Machine Learning

The data was partitioned into a training, validation, and test set.

Two Machine Learning models were used to predict client subscription:

- Logistic Regression

- Random Forest

Because the data set is overwhelmingly skewed towards clients not subscribing, 
the statistical Kappa value helps determine which model will give us best 
results.


## Model Performance Evalution

The Kappa value for the Random Forest model is higher than Logistic Regression 
so the former will be the model of choice for the test data.

Since there is a high volume of rejection, the accuracy of a prediction for 
rejection is overwhelmingly high. Yet we are interested in 

## Recommendations

1. Reach out to subscribers from previous campaign as more than 50% of 
  subscribers will re-subscribe.

2. Contact clients via their cellular phone as they are more than twice as 
  likely to subscribe compared to clients contacted via a regular telephone.

3. Clients are more likely to subscribe when the country's Employment 
  Variation Rate is -1.8. 

