---
title: "Banking Capstone - Data Wrangling and Stastical Analysis"
author: "Demetri Lee"
date: "December 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(GGally)
library(summarytools)
knitr::opts_chunk$set(echo = TRUE)
```

## Working to predict the success of bank marketing

I'm working on predicting the success of telemarketing calls for a bank that's 
selling long-term deposits. A Portuguese retail bank was addressed, with data 
collected from 2008 to 2013, thus including the effects of the recent 
financial crisis of 2008. With 21 variables available for analysis, I hope to 
derive knowledge for a model that would confirm how to better focus resources 
towards clients with a high chance of agreeing to registering for extra 
banking services. 

### This is the data dictionary provided by the respective retail bank:

**Bank client data**

1. age: numeric
2. job: type of job (categorical: admin, blue-collar, entrepreneur,
housemaid, management, retired, self-employed, services, student,
technician, unemployed, unknown)
3. marital: marital status (categorical: divorced, married, single,
unknown; note: 'divorced' means divorced or widowed)
4. education: (categorical: basic.4y, basic.6y, basic.9y, high.school,
illiterate, professional.course, university.degree, unknown)
5. default: has credit in default? (categorical: no, yes, unknown)
6. housing: has housing loan? (categorical: no, yes, unknown)
7. loan: has personal loan? (categorical: no, yes, unknown)

**Related with the last contact of the current campaign**

8. contact: contact communication type (categorical: cellular, telephone) 
9. month: last contact month of year (categorical: jan', 'feb', 'mar', ..., 
'nov', 'dec')
10. day_of_week: last contact day of the week (categorical: mon, tue, wed,
thu, fri)
11. duration: last contact duration, in seconds (numeric). 
Important note: this attribute highly affects the output target (e.g., if 
duration=0 then y='no'). Yet, the duration is not known before a call is 
performed. Also, after the end of the call y is obviously known. Thus, this 
input should only be included for benchmark purposes and should be discarded 
if the intention is to have a realistic predictive model.

**Other attributes**

12. campaign: number of contacts performed during this campaign and for this 
client (numeric, includes last contact)
13. pdays: number of days that passed by after the client was last contacted 
from a previous campaign (numeric; 999 means client was not previously contacted)
14. previous: number of contacts performed before this campaign and for this 
client (numeric)
15. poutcome: outcome of the previous marketing campaign (categorical: 
failure, nonexistent, success)

**Social and economic context attributes**

16. emp.var.rate: employment variation rate - quarterly indicator (numeric)
17. cons.price.idx: consumer price index - monthly indicator (numeric) 
18. cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
19. euribor3m: euribor 3 month rate - daily indicator (numeric)
20. nr.employed: number of employees - quarterly indicator (numeric)

**Output variable (desired target)**

21. y: has the client subscribed a term deposit? (binary: yes, no)

### Data Preparation

Let's start by reading in the dataset with values delimited by semi-colons along 
with designated column types as indicated by the data dictionary.

```{r Data_Prep}
bankfull <- read_delim(file = "bank-additional-full.csv", delim = ";", col_types = "iccccccccciiiicdddddc")

bankfull$job <- as.factor(bankfull$job)
bankfull$marital <- as.factor(bankfull$marital)
bankfull$education <- as.factor(bankfull$education)
bankfull$default <- as.factor(bankfull$default)
bankfull$housing <- as.factor(bankfull$housing)
bankfull$loan <- as.factor(bankfull$loan)
bankfull$contact <- as.factor(bankfull$contact)
bankfull$month <- as.factor(bankfull$month)
bankfull$day_of_week <- as.factor(bankfull$day_of_week)
bankfull$poutcome <- as.factor(bankfull$poutcome)
bankfull$y <- as.factor(bankfull$y)

summary(bankfull)
```

Let's rename variables to be more descriptive and avoid use of periods.

```{r}
bankfull <- rename(bankfull, edu_lvl = "education")
bankfull <- rename(bankfull, cred_default = "default")
bankfull <- rename(bankfull, mortgage = "housing")
bankfull <- rename(bankfull, contact_typ = "contact")
bankfull <- rename(bankfull, duration_sec = "duration")
bankfull <- rename(bankfull, contact_cnt = "campaign")
bankfull <- rename(bankfull, days_passed = "pdays")
bankfull <- rename(bankfull, prev_contact_cnt = "previous")
bankfull <- rename(bankfull, prev_outcome = "poutcome")
bankfull <- rename(bankfull, emp_var_rate = "emp.var.rate")
bankfull <- rename(bankfull, cons_price_idx = "cons.price.idx")
bankfull <- rename(bankfull, cons_conf_idx = "cons.conf.idx")
bankfull <- rename(bankfull, num_employed = "nr.employed")
```

The goal is to determine which factors lead to a subscription so I'll rename 
the dependent variable `y` to `subscribed`, which is more descriptive.

```{r}
bankfull <- rename(bankfull, subscribed = y)
#bankfull$subscribed <- factor(bankfull$subscribed, levels = c("yes", "no"))
```

Replace all periods ('.') in values with underscores ('_') for legibility.

```{r}
summary(bankfull$job) # job types before replacement 
bankfull$job <- gsub("\\.", "", bankfull$job)
bankfull$job <- as.factor(bankfull$job)
summary(bankfull$job) # job types after replacement

summary(bankfull$edu_lvl) # education levels before replacement
bankfull$edu_lvl <- gsub("\\.", "_", bankfull$edu_lvl)
bankfull$edu_lvl <- as.factor(bankfull$edu_lvl)
summary(bankfull$edu_lvl) # education levels after replacement
```

Take a quick check of the dataset with glimpse.

```{r}
glimpse(bankfull)
```

### Exploring Data with Plots

**Looking at Campaign results**

A cursory look at the `subscribe` variable will show most customers do not sign 
up for the extra service so the results are very skewed.

```{r}
ggplot(bankfull, aes(x = subscribed, fill = subscribed)) + geom_bar() + 
  labs(title = "Subscription Counts")
```

The proportion of `no` to `yes` in a table:

```{r}
prop.table(table(bankfull$subscribed))
```

When looking for clients that have subscribed to a previous campaign there are 
too many `nonexistent` values.

```{r}
ggplot(bankfull, aes(prev_outcome, fill=prev_outcome)) + geom_bar() + 
  labs(title="Subscription Count from a Previous Campaign")
```

When comparing clients that subscribed to the previous campaign and the 
current campaign, the `nonexistent` values become distracting.

```{r}
ggplot(bankfull, aes(x=prev_outcome, fill=subscribed)) + geom_bar()
```

With the `nonexistent` values removed, notably more than half the clients can 
be seen as having subscribed again.

```{r}
ctable(bankfull$prev_outcome, bankfull$subscribed)

bankfull_prev <- filter(bankfull, prev_outcome != "nonexistent")
ggplot(bankfull_prev, aes(x=prev_outcome, fill=subscribed)) + 
  geom_bar(position="fill") + 
  labs(title = "Proportion of subscriptions from previous campaign")
```


**Looking at Client Data**

Basic client data such as client age, marital status, education level or job 
type do not seem to show any strong trends.

```{r}
ggplot(bankfull, aes(x = age, fill = subscribed)) + 
  geom_histogram(binwidth=10) + 
  labs(title = "Subscriptions per Age Group")

ggplot(bankfull, aes(x = marital, fill = subscribed)) + 
  geom_bar() + 
  labs(title = "Subscriptions per Marital Status")

ggplot(bankfull, aes(x = edu_lvl, fill = subscribed)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Subscriptions per Education Level")

ggplot(bankfull, aes(x = job, fill = subscribed)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Subscriptions per Job Type")

```

Further review of clients' financial records (credit in default, mortgage, or 
personal loan) also do not express any notable subscription trends.

```{r}
#ggpairs(bankfull, columns=5:7)

ggplot(bankfull, aes(x = cred_default, fill = subscribed)) + 
  geom_bar() + 
  labs(title = "Subscriptions relative to Clients who Default on Credit")

ggplot(bankfull, aes(x = mortgage, fill = subscribed)) + 
  geom_bar() + 
  labs(title = "Subscriptions relative to Clients with a Mortgage")

ggplot(bankfull, aes(x = loan, fill = subscribed)) + 
  geom_bar() + 
  labs(title = "Susbscriptions relative to Clients with an Active Loan")

```

**Last Contact Data**

Last contact data will include details about the last contact with the client. 
Variables include type of contact, the day and month of contact, and the 
duration of the last contact.

Regarding last contact data, it appears customers are more inclined to 
subscribe when contacted via cellular compared to telephone.

```{r}
ggplot(bankfull, aes(x=contact_typ, fill=subscribed)) + geom_bar()
ctable(bankfull$contact_typ, bankfull$subscribed)
ggplot(bankfull, aes(x=subscribed, y=euribor3m, color=subscribed)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

When looking at which month a client is most likely to subscribe, a regular 
bar graph by count makes it difficult. At best it looks like *May* has the 
most subscribers, but that doesn't mean it's the most likely.

```{r}
ggplot(bankfull, aes(x=month, fill=subscribed)) + geom_bar()
```

If we stack each month as relative proportions we can see customers are 
more receptive to subscribing in *March* followed by *December* before 
*September*.

```{r}
ggplot(bankfull, aes(x=month, fill=subscribed)) + geom_bar(position="fill")
```

Here is a cross-tabulation of the monthly subscriptions with explicit results:
```{r}
ctable(bankfull$month, bankfull$subscribed)
```

*March* may have the most percentage of successful subscriptions but the 
number of client engagements is too low to be of significance. The most 
subscriptions do occur in *May* and other summer months, *June, July, and 
August* as confirmed by the frequency chart. Winter months *January* and 
*February* are absent from the dataset.

```{r}
monthsCampaign <- bankfull %>% filter(subscribed == "yes") %>% pull(month)
freq(monthsCampaign)
```

The best day of the week for subscriptions appear to be *Monday* and 
*Thursday*.
```{r}
ctable(bankfull$day_of_week, bankfull$subscribed)
```

But when we integrate weekdays with the monthly values, the second column, 
*Wednesday*, becomes a fairly good day for acquiring subscriptions.

```{r}
ggplot(bankfull, aes(x=day_of_week, fill=subscribed)) + 
  geom_bar(position="fill") +
  facet_wrap(vars(month))
```

The variable `duration` refers to the length of time in seconds spent with the 
customer at the last point of contact. Important note: this attribute highly 
affects the output target (e.g., if duration=0 then subscribed="no"). Yet, 
the duration is not known before a call is performed. Also, after the end of 
the call _subscribed_ is obviously known. Thus, this input should only be 
included for benchmark purposes and should be discarded if the intention is to 
have a realistic predictive model.

```{r}
descr(bankfull$duration_sec)

ggplot(bankfull, aes(x=duration_sec, fill=subscribed)) + geom_histogram(bins=10)

ggplot(bankfull, aes(x=subscribed, y=duration_sec, fill=subscribed)) + 
  geom_boxplot(outlier.shape=16, alpha=.5)

```

There's a slight overlap in the Interquaritle Range of subscribers with the 
non-subscribers. That is, there's a cut-off from the Q3 for non-subscribers 
and Q1 of subscribers.

```{r}
subscribedYES <- filter(bankfull, subscribed=="yes") %>% select(duration_sec)
descr(subscribedYES)

subscribedNO <- filter(bankfull, subscribed=="no") %>% select(duration_sec)
descr(subscribedNO)
```


The following charts affirms the chance of a susbscription occuring decreases 
with each additional contact towards a customer.

```{r}
ggplot(bankfull, aes(x=contact_cnt, fill=subscribed)) + geom_bar(position="fill")
maximum_contact_cnt <- max(bankfull %>% filter(subscribed=="yes") %>% 
  select(contact_cnt))
maximum_contact_cnt
```

As such, the bar graph indicates clients will not subscribe if contacted by 
more than the maximum count value, `r maximum_contact_cnt`.


The following density plot is for the number of days that passed from the previous campaign to the last day of contact. 

```{r}
ggplot(bankfull, aes(x=days_passed)) + geom_density() + facet_wrap(~subscribed)
ctable(bankfull$days_passed, bankfull$subscribed)

```

Since we know a value of 999 means a client was not previously contacted, 
filtering out `days_passed` = 999 should give better insight over days passed 
for clients that were contacted.

```{r}
daysPassedMOD <- filter(bankfull, days_passed!=999)
ggplot(daysPassedMOD, aes(x=subscribed, y=days_passed)) + geom_boxplot()

```


**Socioeconomic Data**

While generating a matrix of plots between the socioeconomic data, there looks 
to be a high correlation in the following 2 pairs:

- euribor3m vs employment variation rate
- euribor3m vs number employed at bank

```{r}
# From emp_var_rate to num_employed

ggpairs(bankfull, columns=16:20, progress=FALSE)

```

**What if I filter out the subscriptions?**

Deep correlation between Employee Variation Rate and Euribor at 3 months but 
doesn't seem like it affects client subscriptions.

```{r correlation}
ggplot(bankfull, aes(emp_var_rate, euribor3m)) + geom_point() + facet_wrap(~subscribed) + geom_smooth(method=lm)
```

Running a scatterplot for the Employment Variation Rate versus the 3 month 
Euribor reiterates the correlation.

```{r}
ggplot(bankfull, aes(x=emp_var_rate, y=euribor3m)) + 
  geom_point(aes(color = factor(subscribed)), alpha = 0.50, position="jitter") + 
  labs(title = "Euribor (3 months) vs Employment Variation Rate (quarterly)")

ggplot(bankfull, aes(x=cons_conf_idx, y=cons_price_idx)) + 
  geom_point(aes(color=factor(subscribed), shape=factor(subscribed)), alpha=0.50, position="jitter") + 
  labs(title = "Consumer Price Index vs Consumer Confidence Index")
```

Using box-plots used for the employment variable rate and the 3 month Euribor 
rating with respect to subscriptions continue to look similar.

```{r}
ggplot(bankfull, aes(x=subscribed, y=emp_var_rate, fill=subscribed)) + geom_boxplot()
ggplot(bankfull, aes(x=subscribed, y=euribor3m, fill=subscribed)) + geom_boxplot()
```


For this exploratory data analysis phase, here are a few conclusions:

1. Reach out to subscribers from previous campaign as more than 50% of 
  subscribers will re-subscribe.
2. Contact clients via their mobile phone as the success rate is more than 
  twice than clients contacted via a regular telephone.
3. It is advised to close a subscription by the 11th contact because there is 
  a strong chance of rejection.
4. The length of the call should be more than 274 seconds or 4 min, 34 sec.

_Write the revised dataset into a cleaned-up file._

```{r conclusion}
write_csv(bankfull, path="bankfull_clean.csv", col_names=TRUE)
```
