---
title: "Banking Capstone Data Wrangled"
author: "Demetri Lee"
date: "December 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Trying to predict the success of bank telemarketing

I'm trying to predict the success of telemarketing calls for a bank that's 
selling long-term deposits. A Portuguese retail bank was addressed, with data 
collected from 2008 to 2013, thus including the effects of the recent 
financial crisis of 2008. With 21 features available for analysis, I hope to 
derive knowledge for a model that would confirm how to better focus resources 
towards clients with a high chance of agreeing to registering for extra 
banking services. 

**Read in the dataset with values delimited by semi-colons along with 
designated column types.**
```{r}
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
```

Rename variables to be more descriptive and avoid use of periods (".")
```{r}
bankfull <- rename(bankfull, edu_lvl = "education")
bankfull <- rename(bankfull, cred_def = "default")
bankfull <- rename(bankfull, duration_sec = "duration")
bankfull <- rename(bankfull, contact_typ = "contact")
bankfull <- rename(bankfull, contact_cnt = "campaign")
bankfull <- rename(bankfull, days_passed = "pdays")
bankfull <- rename(bankfull, prev_contact_cnt = "previous")
bankfull <- rename(bankfull, prev_outcome = "poutcome")
bankfull <- rename(bankfull, emp_var_rate = "emp.var.rate")
bankfull <- rename(bankfull, cons_price_idx = "cons.price.idx")
bankfull <- rename(bankfull, cons_conf_idx = "cons.conf.idx")
bankfull <- rename(bankfull, num_employed = "nr.employed")
```

The goal is to determine what factors lead to a subscription so I'll rename 
the dependent variable to something more descriptive than simply, 'y'.
```{r}
bankfull <- rename(bankfull, subscribed = y)
```

Replace all periods ('.') with underscores ('_') for legibility
```{r}
etemp <- bankfull$edu_lvl
bankfull$edu_lvl <- gsub("\\.", "_", etemp)
```

There are only 2 possible values for the dependent variable, "subscribed". As
such I'll convert the variable to type factor.
```{r}
bankfull$subscribed <- as.factor(bankfull$subscribed)
```

The variable prev_outcome contains the result of the previous marketing 
campaign with values "failure", "nonexistent", and "success".
```{r}
bankfull$prev_outcome <- as.factor(bankfull$prev_outcome)
```

Take a glimpse at the dataset
```{r}
glimpse(bankfull)
```

### Including Plots

A cursory look at the subscribed variable will show most customers do not sign 
up for the extra service.
```{r}
prop.table(table(bankfull$subscribed))
ggplot(bankfull, aes(x = subscribed)) + geom_bar() + labs(title = "How many clients subscribed?")
```

How many clients from the previous campaign also signed up for a term deposit 
this time around? Similar to looking for a repeat customer.
```{r}
xtabs(~prev_outcome+subscribed, bankfull)
ggplot(bankfull, aes(prev_outcome)) + geom_bar() + labs(title = "How many clients subscribed to a previous campaign?")
plot(xtabs(~prev_outcome+subscribed, bankfull), 
     main="Previous Campaign Outcome vs Current Subscrption")
```

** Why did the previous campaign contain so many non-existant responses? **
```{r}
ggplot(bankfull, aes(x = subscribed, fill = prev_outcome)) + 
  geom_bar() + 
  labs(title = "How many customers from previous campaign also bought into current subscription?")
```

** Is there any positive correlation between having a default on credit or having a loan and a subscription? ** 
```{r}
ggplot(bankfull, aes(x = subscribed, fill = cred_def)) + 
  geom_bar() + 
  labs(title = "Do customers with credit in default tend to describe?")

ggplot(bankfull, aes(x = subscribed, fill = housing)) + 
  geom_bar() + 
  labs(title = "Do customers with a housing loan tend to subscribe?")

ggplot(bankfull, aes(x = subscribed, fill = loan)) + 
  geom_bar() + 
  labs(title = "Do customers with a personal loan tend to subscribe?")

ggplot(bankfull, aes(x = housing, fill = subscribed)) + 
  geom_bar() + 
  facet_wrap(~loan)

```

__ It seems those with defaults on credit or personal loans are not lightly to 
subscribe but those with a housing loan may subscribe.__

** Let's look at the age of our customers **

```{r}
ggplot(bankfull, aes(x = age, fill = subscribed)) + 
  geom_histogram(binwidth = 5) + 
  labs(title = "Age distrubtion of current bank clients and those who subscribed") 
```

```{r}
ggplot(bankfull, aes(x = subscribed, y = age)) + 
  geom_boxplot() + 
  labs(title = "Subscription rates by age")
```



The variable __duration__ refers to the length of time the customer was last 
contacted, in seconds. Important note:  this attribute highly affects the 
output target (e.g., if duration=0 then subscribed="no"). Yet, the duration 
is not known before a call is performed. Also, after the end of the call 
__subscribed__ is obviously known. Thus, this input should only be included 
for benchmark purposes and should be discarded if the intention is to have a 
realistic predictive model.

```{r}
bankfull <- bankfull %>% mutate(duration_min = duration_sec/60)
summary(bankfull$duration_min)
boxplot(bankfull$duration_min)
#ggplot(bankfull, aes(x=duration)) + geom_bar()
```


```{r}
plot(density(bankfull$age), main="Density of Customer Age")
ggplot(bankfull, aes(x=age)) + geom_bar()
ggplot(bankfull, aes(x=age)) + geom_histogram(binwidth = 1)
```

```{r fig.cap = "Contact Type vs Subscribed"}
#count(bankfull, vars = day_of_week)
#ggplot(bankfull, aes(x=subscribed, fill = contact_typ)) + geom_bar()
ggplot(bankfull, aes(x=contact_typ, fill = subscribed)) + geom_bar()
#try geom_col()
count(bankfull)

```

Outcome of previous campaign

```{r}
ggplot(bankfull, aes(x = prev_outcome)) + geom_bar()
summary(bankfull$prev_outcome)
```


How many of the same customers subscribed to a long-term deposit service versus those who have not?

```{r echo=TRUE, fig.cap = "Only 11.3% of customers subscribed while an overwhelming 88.7% customers declined"}
reject_cnt <- tally(bankfull, subscribed == "no")
reject_pct <- reject_cnt/nrow(bankfull)*100
accept_cnt <- tally(bankfull, subscribed == "yes")
accept_pct <- accept_cnt/nrow(bankfull)*100
summary(bankfull$subscribed)
ggplot(bankfull, aes(x = subscribed)) + geom_bar()

print(accept_pct)
#print(reject_pct, "% would reject.")

```
test = `r reject_cnt + 20`



How did a previous campaign perform?

```{r fig.cap = "failure = 10.3% vs success = 3.3% while remaining 35,563 were non-existent"}
ggplot(bankfull, aes(x = prev_outcome)) + geom_bar()
```


Write the revised dataset into a conclusive file.
```{r}
write_csv(bankfull, path="bankfull_clean.csv")
```

