---
title: "Banking Capstone Wrangled"
author: "Demetri Lee"
date: "December 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Trying to predict the success of bank telemarketing

I'm trying to predict the success of telemarketing calls for a bank that's 
selling long-term deposits. A Portuguese retail bank was addressed, with data 
collected from 2008 to 2013, thus including the effects of the recent 
financial crisis of 2008. With 21 features available for analysis, I hope to 
derive knowledge for a model that would confirm how to better focus resources 
towards clients with a high chance of agreeing to registering for extra 
banking services. 

**Read in the dataset with values delimited by semi-colons along with 
designated column types.**

```{r}
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
```

Rename variables to be more descriptive and avoid use of periods (".")

```{r}
bankfull <- rename(bankfull, edu_lvl = "education")
bankfull <- rename(bankfull, contact_typ = "contact")
bankfull <- rename(bankfull, contact_cnt = "campaign")
bankfull <- rename(bankfull, days_passed = "pdays")
bankfull <- rename(bankfull, prev_contact_cnt = "previous")
bankfull <- rename(bankfull, prev_outcome = "poutcome")
bankfull <- rename(bankfull, emp_var_rate = "emp.var.rate")
bankfull <- rename(bankfull, cons_price_idx = "cons.price.idx")
bankfull <- rename(bankfull, cons_conf_idx = "cons.conf.idx")
bankfull <- rename(bankfull, num_employed = "nr.employed")
```

The goal is to determine what factors lead to a subscription so I'll rename 
the dependent variable to something more descriptive than simply, 'y'.
```{r}
bankfull <- rename(bankfull, subscribed = y)
```


Replace all periods ('.') with underscores ('_') for legibility

```{r}
etemp <- bankfull$edu_lvl
bankfull$edu_lvl <- gsub("\\.", "_", etemp)
```

Take a glimpse at the dataset

```{r}
glimpse(bankfull)
```

### Including Plots

```{r}
ggplot(bankfull, aes(x=age)) + geom_histogram(binwidth = 5)
```
```{r}
count(bankfull, vars = day_of_week)
ggplot(bankfull, aes(x=day_of_week)) + geom_bar()
```


How many of the same customers subscribed to a long-term deposit service versus those who have not?

```{r echo=TRUE, fig.cap = "Only 11.3% of customers subscribed while an overwhelming 88.7% customers declined"}
reject_cnt <- tally(bankfull, subscribed == "no")
accept_cnt <- tally(bankfull, subscribed == "yes")
ggplot(bankfull, aes(x = subscribed)) + geom_bar()

#print("Only ", accept_cnt/nrows(bankfull), "%")

```
test = `r reject_cnt + 20`



How did a previous campaign perform?

```{r fig.cap = "failure = 10.3% vs success = 3.3% while remaining 35,563 were non-existent"}
ggplot(bankfull, aes(x = prev_outcome)) + geom_bar()
```


