---
title: "Bank Marketing Success"
author: "Demetri Lee"
date: "9 April 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(tidyverse)
library(summarytools)

#bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "ifffffffffiiiifdddddf")

colnames(bankfull) <- c("age", "job", "marital", "edu_lvl", "cred_default", "mortgage", "loan",
                        "contact_typ", "month", "day_of_week", "duration_sec", 
                        "contact_cnt", "days_passed", "prev_contact_cnt", "prev_outcome",
                        "emp_var_rate", "cons_price_idx", "cons_conf_idx", "euribor3m", "num_employed",
                        "subscribed")
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

A Portuguese bank uses telemarketing campaigns to engage their clients.

Each client engagement has been recorded for evaluation.

Such data can be used to extract insight for making better campaign decisions.


## Objective

Marketing campaigns generally do not discriminate which clients are targeted.

Therefore, some campaigns are irrelevant to some clients.

As such, unneccessary resources may be spent on clients that lead to rejection. 

The goal is to predict whether a client will subscribe to a term deposit.


## Data

Interactions with over 41,000 clients were recorded from May 2008 to Nov 2010.

Each record contains 21 different attributes for evaluation.

The attribute of interest, `subscribed`, will indicate a successful 
subscription from a client.


## Exploratory Data Analysis

At first glance there is an overwhelming amount of clients who choose not to 
subscribe to a term deposit.

```{r subscribed}
ggplot(bankfull, aes(x = subscribed, fill = subscribed)) + geom_bar() + 
  xlab("subscribed") + 
  ylab("count")
```


## Exploratory Data Analysis

But looking at subscriptions from a previous campaign tells us more than 
half the subscribers will re-subscribe.

```{r prev_outcome}
bankfull_prev <- filter(bankfull, prev_outcome != "nonexistent")
ggplot(bankfull_prev, aes(x=prev_outcome, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("previous campaign") + 
  ylab("ratio")
```


## Contact Type

It was also observed contacting clients via cellular phone yields at least 2 
times more success than contacting a client through their landline telephone.

```{r contact_typ}
ggplot(bankfull, aes(x=contact_typ, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("method of contact") + 
  ylab("ratio")
```


## Employment Variation Rate

Logistic Regression indicates there is a fair chance of success when the 
country's Employment Variation Rate is -1.8.

```{r emp_var_rate}
ggplot(bankfull, aes(x=subscribed, y=emp_var_rate, fill=subscribed)) +
  geom_boxplot() + geom_jitter(width=0.2) + 
  xlab("subscribed")
  ylab("employment variation rate")
```


## Machine Learning

The data was partitioned into a training, validation, and test set.

1. A predictive model is created based on client responses in the training set.

2. The same model is used to predict client response in the validation set.

3. Finally the test set is used to evaluate the model created from the training 
  set.


## Predictive Modeling

Two Machine Learning models were used to predict client subscription:

* Logistic Regression

* Random Forest

Because the data set is overwhelmingly skewed towards clients not subscribing, 
the statistical Kappa value helps determine which model will give us the best 
results.


## Model Performance Evalution

The Kappa value for the Random Forest model is higher than the Logistic 
Regression model so the former will be the model of choice for the test data.

Since there is a high volume of rejection, the accuracy of a prediction for 
rejection is overwhelmingly high. Yet we are interested in successful 
subscriptions, which are true positive results.

Sensitivity vs. Specificity

## Recommendations

1. Reach out to subscribers from previous campaign as more than 50% of 
  subscribers will re-subscribe.

2. Contact clients via their cellular phone as they are more than twice as 
  likely to subscribe compared to clients contacted via a regular telephone.

3. Clients are more likely to subscribe when the country's Employment 
  Variation Rate is -1.8. 

